---
title: "Water Quality in North Carolina: Nutrients and Specific Conductance" 
author: "Peggy Chen"
date: "2024-06-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

Using the USGS "dataRetrieval" package to plot the frequency of measuring certain chemicals across North Carolina. 

Paramter codes, or PCodes, are dependent on the chemical or physical factor being tested. 
For a list of the codes, go to: https://help.waterdata.usgs.gov/code/parameter_cd_query?fmt=rdb&group_cd=NUT&inline=true 

Need help? Type into the console: HASP::explore_site() 

### Installing the USGS HASP package 
```{r}
# First install the remotes package 
# install.packages("remotes") 

# Install HASP package 
# remotes::install_gitlab("water/stats/hasp", 
                        # host = "code.usgs.gov", 
                        # build_opts = c("--no-resave-data","--no-manual"), 
                        # build_vignettes = TRUE, 
                        # dependencies = TRUE) 
```


### Loading and prepping data 

```{r}
library(tidyverse) 
library(remotes)
library(HASP)
library(dataRetrieval)
library(dplyr)
library(ggplot2)
library(sf)
library(olsrr)
```

### Plotting total nitrogen testing sites across NC 

Note that total nitrogen in this plot is defined as nitrate, nitrite, ammonia, organic-N, and is tested in water, unfiltered, and measured in milligrams per liter. 
```{r}
pcode <- readNWISpCode("all")

phosCds <- pcode[grep("specific conductance",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

names(phosCds)
```


```{r}
# What sites measured nitrogen in North Carolina? 
nitrogen_sites <- whatNWISsites(stateCd = "NC", 
                          parameterCd = "00600")

# Identify these sites 
names(nitrogen_sites)

# Plot the locations of these sites 
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)

sf_nitrogen <- st_as_sf(nitrogen_sites, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)

# Bad map :( 
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_nitrogen) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Total Nitrogen in Water across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021")

```

### Plotting ammonia testing sites across NC 

```{r}
ammonia_sites <- whatNWISsites(stateCd = "NC", 
                          parameterCd = "00608")
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)
sf_ammonia <- st_as_sf(ammonia_sites, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_ammonia) + 
  xlab(NULL) +
  ylab(NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Ammonia in Water across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
```

### Plotting phosphate freq across state 

```{r}
phosphate_site <- whatNWISsites(stateCd = "NC", 
                               parameterCd = "00650")
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)
sf_phosphate <- st_as_sf(phosphate_site, 
                       coords = c("dec_long_va", "dec_lat_va"),
                       crs = 4269)
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_phosphate) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Phosphate in Water across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
``` 

### Field Specific Conductance freq across NC Map  

Specific conductance, water, unfiltered, field, microsiemens per centimeter at 25 degrees. 

```{r}
conductivity_sites <- whatNWISsites(stateCd = "NC", 
                          parameterCd = "00094")
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)
sf_conductivity <- st_as_sf(conductivity_sites, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_conductivity) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Field Specific Conductance across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
```

### Non-field SC in NC Map  

Specific conductance, water, unfiltered, microsiemens per centimeter at 25 degrees Celsius. 

```{r}
conductivity_sites_2 <- whatNWISsites(stateCd = "NC", 
                          parameterCd = "00095")
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)
sf_conductivity_2 <- st_as_sf(conductivity_sites_2, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_conductivity_2) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Specific Conductance across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
```

### Another SC across NC Map 

Specific conductance, water, unfiltered, laboratory, microsiemens per centimeter at 25 degrees Celsius	USGS use and no use by EPA. Total	25 deg C. Specific conductance uS/cm. 

```{r}
conductivity_sites_3 <- whatNWISsites(stateCd = "NC", 
                          parameterCd = "90095")
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE),
                crs = 4269)
sf_conductivity_3 <- st_as_sf(conductivity_sites_3, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)
ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_conductivity_3) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "USGS Locations that Meausure Specific Conductance across North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
```

### Filtered Data Example 

This example filters the points so that rather than displaying all locations where 
nitrogen is tested, it only finds sites with more than 100 measurements and at least 
5 measurements per year 

```{r}
# Assign codes and pull data 
pCode <- c("00600")
phNC <- whatNWISdata(stateCd="NC", 
                     parameterCd=pCode)

# Filter data acc. to freq and measurement count 
phNC_1 <- phNC %>% 
  filter(count_nu > 100) %>% # At least 100 measurments 
  mutate(period = as.Date(end_date) - as.Date(begin_date)) %>%
  filter(period > 5*365) # At least 5x/year 

nitrogen_NC_data <- readNWISqw(siteNumbers = phNC_1$site_no,
                         parameterCd = pCode)

# Let’s look at the maximum measured value, and number of samples 
nitrogen_summary <- nitrogen_NC_data %>% 
  group_by(site_no) %>% 
  summarize(max = max(result_va, na.rm = TRUE),
            count = n()) %>% 
  ungroup() %>% 
  left_join(attr(nitrogen_NC_data, "siteInfo"), 
            by = "site_no")

# Prep data to map >:( 
sf_nitrogen <- st_as_sf(nitrogen_summary, 
                       coords = c("dec_long_va", "dec_lat_va"),
                       crs = 4269)

# Map 
plot_1 <- ggplot() +
  geom_sf(data = usa[ usa$ID == "north carolina" ,]) +
  geom_sf(data = sf_nitrogen) + 
  xlab(NULL)+
  ylab(NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  labs(title = "Nitrogen in North Carolina", 
       caption = "Laura DeCicco, USGS. 2021") 
plot_1
```

### Plotting over time 

These plots are measurements of specific conductance overtime, and the site codes are obtained 
from the "sf_conductivity" data that can be found in the environment. 

```{r}
site <- "02129000"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point()
plot
```

### Trying more sites 

```{r}
site <- "02092500"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_1 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point()

plot_1

```

```{r}
site <- "335334078352105"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_2 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point()

plot_2
```

```{r}
site <- "02083500"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_3 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point()

plot_3 
```

```{r}
site <- "02091500"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_4 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_4 
```
```{r}
site <- "02089500"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_5 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_5 
```

```{r}
site <- "02105769"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_6 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_6 
```

```{r}
site <- "02105524"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_7 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_7 
```

```{r}
site <- "02098198"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_8 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_8 
```

```{r}
site <- "02129000"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance")) 

plot_9 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_9 
```

```{r}
site <- "03460000"

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance"))

plot_10 <- ggplot(data = qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) + 
  geom_point() 

plot_10 
```

### Calculating Linear Regression and R Squared 

```{r}
site <- "02129000" 

qw_data <- readWQPqw(paste0("USGS-", site),
                     c("Specific conductance")) 

y <- qw_data$ResultMeasureValue 
x <- qw_data$ActivityStartDate

# Estimate the model and same the results in object "ols"
ols <- lm(y ~ x, data = qw_data) 

# Print the result in the console
ols

# Determine accuracy 
summary(ols)$r.squared

# Now graph the regression line 
regression_1 <- ggplot(qw_data, aes(x = ActivityStartDate, y = ResultMeasureValue)) +
  geom_point() + 
  geom_smooth(method = "lm",  fill="#69b3a2", se = TRUE) + 
  labs(title = "Specific Conductance over Time at Pee Dee River in Rockingham, NC", 
       x= "Year", 
       y = "Specific conductance in µS/cm at 25 °C", 
       caption = "USGS Water Quality Data, 2024.") + 
  theme_bw() 

regression_1 
``` 